---
title: "MY ANALYSIS"
author: "MY NAME"
date: "03/05/2022"
output:
  html_document:
    fig_height: 6
    fig_width: 9
  pdf_document: default
  word_document:
    fig_height: 6
    fig_width: 9
---

```{r clean, include=FALSE, purl=FALSE}
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
#gc() #free up memory and report the memory usage.
```

# Title
**TITLE OF MY PROJECT**

Description of the project

# Packages

Load required packages

```{r package setup, echo=FALSE, warning=F}
#install.packages("data.table")
#install.packages("psych")
#install.packages("stats")  
#install.packages("tidyverse") 

library(data.table)
library(psych)
library(stats)
library(tidyverse)

```

# Script settings

Set variables that control the overall behaviour of the script

```{r settings}

mySetting <- 42

```

# Functions

Functions can be used for code that you want to re-use multiple times throughout your script.

```{r general functions}

printHello <- function(argument1Greeting="Hello"){
  
  print(paste0(argument1Greeting,", we want to do things in here many times in the project."))
  
}

```

#Knitr settings

These are settings that control Rmd behaviour.

```{r knitr setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment=NA,prompt=FALSE,cache=FALSE)
#knitr::opts_knit$set(root.dir=myNewWorkingDirectoryPath) #you can set the default working dir for Rmd-files here 

```

# R script export

```{r purl export, include=FALSE, purl=FALSE, eval=FALSE}
#running this will produce an R-script with the same name as the Rmd-file. Used for running the program on computational clusters (Rosalind), when the code just has to be run rather than creating a knitted output.
# knitr::purl(filenameOfThisRmdFile_HasToBeSetSomewhere)
  
```

# Additional R source

You can execute more R code by sourcing in the files

```{r additional source setup, echo=FALSE, warning=F}

#source(file.path(folderpath.scripts,"include","fastqq.R"))

```


# My Analysis part 1

```{r myAnalysis}

printHello(argument1Greeting = "Howdy")
printHello(argument1Greeting = "Greetings")

```




